---
title: "Canopy Height Model"
author: "K Perham"
date: "May 2017"
output: html_document
---
---
title: "Calculating Canopy Height"
author: "K Perham"
date: "May 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
```
```{r include=FALSE}
library(raster)
library(rgdal)
library(dplyr)
```
## Gathering and plotting DTM and DSM files

This exercise to create a Canopy Height Model first calls for the DTM and DSM files to be retrieved and plotted:

```{r echo=FALSE}
lidar_dem <- raster(x="c:/users/kaitlyn/documents/earth-analytics/data/week3/BLDR_LeeHill/pre-flood/lidar/pre_DTM.tif")
```

```{r echo=FALSE}
plot(lidar_dem,
     main="Lidar Digital Elevation Model (DEM)")
```
```{r echo=FALSE}
lidar_dsm <- raster(x="c:/users/kaitlyn/documents/earth-analytics/data/week3/BLDR_LeeHill/pre-flood/lidar/pre_DSM.tif")
```

```{r echo=FALSE}
plot(lidar_dsm,
     main="Lidar Digital Surface Model (DSM)")
```

Now that we have gathered both rasters, the CHM calculation is simple in r:
```{r echo=TRUE}
lidar_chm <- lidar_dsm - lidar_dem
```
The resulting plot looks like this:
```{r echo=FALSE}
plot(lidar_chm,
     main="Lidar Canopy Height Model (CHM)")
```
Breaks can be created to improve the visualization.

```{r echo=FALSE}
plot(lidar_chm,
     breaks = c(0, 2, 10, 20, 30),
     main="Lidar Canopy Height Model",
     col=c("white","brown","springgreen","darkgreen"))
```

And finally, a histogram of the data, eliminating the outliers to get a better idea of the distribution of values.

```{r echo=FALSE}
hist(lidar_chm,
     xlim=c(2,25),
     ylim=c(0,4000),
     main="Histogram of canopy height model differences \nZoomed in to -2 to 2 on the x axis",
     col="springgreen")
```
```{r echo=FALSE}
reclass_df <- c(0, 2, NA,
              2, 4, 1,
             4, 7, 2,
             7, Inf, 3)
reclass_m <- matrix(reclass_df,
                ncol=3,
                byrow=TRUE)
```
We can reclassify the raster into short, medium, and tall trees. It's interesting to see how much more visible human habitation is in this plot, just based on tree height values. 
```{r include=FALSE}
chm_classified <- reclassify(lidar_chm,
                     reclass_m)
chm_classified[chm_classified==0] <- NA
```
```{r echo=FALSE}
plot(chm_classified,
     legend=F,
     col=c("red", "blue", "green"), axes=F,
     main="Classified Canopy Height Model \n short, medium, tall trees")
legend("topright",
       legend = c("short trees", "medium trees", "tall trees"),
       fill = c("red", "blue", "green"),
       border = F,
       bty="n") 
```